N-Body problem

### 2014-10-25 article/article.tex
Fixes typos

This research is aims to provide insight in the approximation methods used for 2-body systems, i.e. how accurate is the simulation. 
In section~\ref{sec:literature} the background of the analytical models used are explained and section~\ref{sec:methodology} introduces the algorithms as well a methods for benchmarking our approximation methods. The results are shown in section~\ref{sec:results} and section~\ref{sec:conclusion} our conclusion is discussed.

@@

\subsection{Newton's laws}
Newton's second law describes that the force on an object is equal to the change of momentum in time. Because we only consider constant mass systems this is equal to the mass times the acceleration of an object. Thus, if the object is accelerating there is an force on our object present. Knowing that the only possible object acting on our object are the other planets in the system we can modify the second law to derive the law of universal gravitation. Let the following equation be the force $F_p$ on object $p$ by object $c$ with mass $m_p,m_c$ respectively.   
\begin{equation}
    \label{eq:newtongravity}
 F_p = G\frac{m_pm_c}{r^2}
\end{equation}
With $r$ as the euclidean distance between object $p,c$. Combining Newton's second law with the law of universal gravitation gives the acceleration $a_p$ of object $p$
\begin{equation}
    \label{eq:newtongravity}
	a_p = G\frac{m_pm_c}{ m_pr^2}
\end{equation}.
The third law states that if there is a force $F_1$ then there also exists a force $F_2$ such that $F_2 = - F_1$.

@@

% Assume that we have an initial velocity vector $\overrightarrow{v_t}$ for the body t with speed in both x and y direction.
% \begin{equation}
% 	\overrightarrow{v_t} = \begin{bmatrix}
% 								x_t \\
% 								y_t
% 							\end{bmatrix}
% \end{equation}
% For example we have a two body system with $p,c$ with an initial velocity vector, respectively, then
% \begin{equation}
% 	v_p=\begin{bmatrix} v_x^p \\ v_y^p \end{bmatrix}, v_c=\begin{bmatrix} v_x^c \\ v_y^c \end{bmatrix}
% \end{equation}
We assume that the whole system has a constant velocity, in our case we assume zero velocity, hence the next equation must be equal to zero
\begin{equation}
	V = \frac{P}{M}	
\end{equation}
with $P$ as the total momentum and $M$ as the total mass of the system. We know that the total momentum is the sum of all forces acted up on and the mass is a constant. Therefore, our zero velocity implies that the total momentum must be equal to zero. Or more formally
\begin{equation}
	P = m_p\overrightarrow{v_p} + m_c\overrightarrow{v_c} = 0.
\end{equation}
As illustrated in figure~\ref{XXXX}(simple two body system with vectors drawn) we can see that each body has several forces acted upon. Our objective is the determine the path of body $p,c$ and as you can see we have a force vector $F_x$ pulling the body $X$ to the center and also our velocity vector $v_x$. The product of those vectors gives us the directional vector of our body $X$. Due to the fact that the two bodies are the only objects in the system acting on eachother we can determine the radius and period of our movement around a (virtual) center of mass. We know for a fact that the two bodies must be the opposites of eachother such that the total momentum is equal to zero. The center of mass can be calculated in reference of one of our bodies.  (bary center)
\begin{equation}
	(m_p + m_c)\overrightarrow{r_{m}} = m_p\overrightarrow{r_p} + m_c\overrightarrow{r_c} 
\end{equation}
With $r_m$ being the vector to the center of mass, the we can rewrite this to calculate $r_m$
\begin{equation}
	\overrightarrow{r_m} = \frac{m_p\overrightarrow{r_p} + m_c\overrightarrow{r_c}}{(m_p + m_c)}
\end{equation}
and we will calculate the $\overrightarrow{r_m}$ with as reference point our body $C$, then the $\overrightarrow{r_c}=0$ because the distance between our body $C$ and $C$ is zero, and the $\overrightarrow{r_p}=R$ with $R$ as the total distance between the two bodies. Then $\overrightarrow{r_m}$ becomes
\begin{equation}
	\overrightarrow{r_m} = \frac{m_p}{M}\overrightarrow{R}
\end{equation}
force times direction (unit vector) see albana notes on slide
Calculate $a_i$ for each body i.

If we want to know the force of body $p$ on body $c$ we can use Newton's law of gravitation, see equation~\ref{eq:newtongravity}.  We can fill in the masses of bodies $p,c$ using the fact that $F_c = m_c a_c$ we obtain
\begin{equation}
    \begin{split}
     a_c m_c = G\frac{m_c m_p}{r^2} \\
     a_c = G\frac{m_p}{r^2} \\
    \end{split}
\end{equation}
Now that we have derived the acceleration in terms of the mass of $p$ and the distance between the bodies $r$, we can use this calculate the acceleration of body $c$. The change of position i.e. distance of body $c$ at time $\delta t$ is given by
\begin{equation}
    \frac{d^2}{dt^2} r_c = G \frac{m_p}{r^2}
\end{equation}
So the new position can be derived by taking the position at time $t=0$ plus the distance traveled during $\delta t$. This can be obtained by taking the double integral of the acceleration. This can be solved analytically when only considering two bodies. If you consider $n \geq 3$ bodies it becomes unsolvable analytically. 
\subsection{Discrete Approximation}
The N-body system for more than 2 bodies can not be solved analytically, hence we must use approximation methods to calculate the solution.
We want to obtain the new positions of our bodies at time $t+1$ by using Newton's second law. We can derive the next position by adding the distance traveled in time $\Delta t$ to the old position at $x_t$, i.e. $x_{t+1} = x_t + v_t*\Delta t$ wit $v_t$ as the speed at time $t$. Nonetheless, we do not know speed $v_t$ but this can be obtained by $v_t = v_t + a_t \Delta t$ and $a_t$ is known for our body. The next section describes an even more accurate approach for $x_t$.

>>>>

This research is aims to provide insight in the approximation methods used for N-body systems, i.e. how accurate is the simulation. 
In section~\ref{sec:literature} the background of the analytical models used are explained and section~\ref{sec:methodology} introduces the algorithms as well a methods for benchmarking our approximation methods. The results are shown in section~\ref{sec:results} and section~\ref{sec:conclusion} our conclusion is discussed.

@@

\subsection{Newton's laws}
Newton's second law describes that the force on an object is equal to the change of momentum in time. Because we only consider constant mass systems this is equal to the mass times the acceleration of an object. Thus, if the object is accelerating there is a force on our object present. Knowing that the only possible object acting on our object are the other planets in the system we can modify the second law to derive the law of universal gravitation. Let the following equation be the force $F_p$ on object $p$ by object $c$ with mass $m_p,m_c$ respectively.   
\begin{equation}
    \label{eq:newtongravity}
 F_p = G\frac{m_pm_c}{r^2}
\end{equation}
With $r$ as the euclidean distance between object $p,c$. Combining Newton's second law with the law of universal gravitation gives the acceleration $a_p$ of object $p$
\begin{equation}
    \label{eq:newtongravity}
	a_p = G\frac{m_pm_c}{ m_pr^2}
\end{equation}.
The third law states that if there is a force $F_1$ then there also exists a force $F_2$ such that $F_2 = - F_1$.

@@

% Assume that we have an initial velocity vector $\overrightarrow{v_t}$ for the body t with speed in both x and y direction.
% \begin{equation}
% 	\overrightarrow{v_t} = \begin{bmatrix}
% 								x_t \\
% 								y_t
% 							\end{bmatrix}
% \end{equation}
% For example we have a two body system with $p,c$ with an initial velocity vector, respectively, then
% \begin{equation}
% 	v_p=\begin{bmatrix} v_x^p \\ v_y^p \end{bmatrix}, v_c=\begin{bmatrix} v_x^c \\ v_y^c \end{bmatrix}
% \end{equation}
We assume that the whole system has a constant velocity, in our case we assume zero velocity, hence the next equation must be equal to zero
\begin{equation}
	V = \frac{P}{M}	
\end{equation}
with $P$ as the total momentum and $M$ as the total mass of the system. We know that the total momentum is the sum of all forces acted up on and the mass is a constant. Therefore, our zero velocity implies that the total momentum must be equal to zero. Or more formally
\begin{equation}
	P = m_p\overrightarrow{v_p} + m_c\overrightarrow{v_c} = 0.
\end{equation}
As illustrated in figure~\ref{XXXX}(simple two body system with vectors drawn) we can see that each body has several forces acted upon. Our objective is the determine the path of body $p,c$ and as you can see we have a force vector $F_x$ pulling the body $X$ to the center and also our velocity vector $v_x$. The product of those vectors gives us the directional vector of our body $X$. Due to the fact that the two bodies are the only objects in the system acting on each other we can determine the radius and period of our movement around a (virtual) center of mass. We know for a fact that the two bodies must be the opposites of each other such that the total momentum is equal to zero. The center of mass can be calculated in reference of one of our bodies.  (Bary center)
\begin{equation}
	(m_p + m_c)\overrightarrow{r_{m}} = m_p\overrightarrow{r_p} + m_c\overrightarrow{r_c} 
\end{equation}
With $r_m$ being the vector to the center of mass, the we can rewrite this to calculate $r_m$
\begin{equation}
	\overrightarrow{r_m} = \frac{m_p\overrightarrow{r_p} + m_c\overrightarrow{r_c}}{(m_p + m_c)}
\end{equation}
and we will calculate the $\overrightarrow{r_m}$ with as reference point our body $C$, then the $\overrightarrow{r_c}=0$ because the distance between our body $C$ and $C$ is zero, and the $\overrightarrow{r_p}=R$ with $R$ as the total distance between the two bodies. Then $\overrightarrow{r_m}$ becomes
\begin{equation}
	\overrightarrow{r_m} = \frac{m_p}{M}\overrightarrow{R}
\end{equation}

If we want to know the force of body $p$ on body $c$ we can use Newton's law of gravitation, see equation~\ref{eq:newtongravity}.  We can fill in the masses of bodies $p,c$ using the fact that $F_c = m_c a_c$ we obtain
\begin{equation}
    \begin{split}
     a_c m_c = G\frac{m_c m_p}{r^2} \\
     a_c = G\frac{m_p}{r^2} \\
    \end{split}
\end{equation}
Now that we have derived the acceleration in terms of the mass of $p$ and the distance between the bodies $r$, we can use this calculate the acceleration of body $c$. The change of position i.e. distance of body $c$ at time $\delta t$ is given by
\begin{equation}
    \frac{d^2}{dt^2} r_c = G \frac{m_p}{r^2}
\end{equation}
So the new position can be derived by taking the position at time $t=0$ plus the distance traveled during $\delta t$. This can be obtained by taking the double integral of the acceleration. This can be solved analytically when only considering two bodies. If you consider $n \geq 3$ bodies it becomes unsolvable analytically. 
\subsection{Discrete Approximation}
The N-body system for more than 2 bodies can not be solved analytically, hence we must use approximation methods to calculate the solution.
We want to obtain the new positions of our bodies at time $t+1$ by using Newton's second law. We can derive the next position by adding the distance traveled in time $\Delta t$ to the old position at $x_t$, i.e. $x_{t+1} = x_t + v_t*\Delta t$ with $v_t$ as the speed at time $t$. Nonetheless, we do not know speed $v_t$ but this can be obtained by $v_t = v_t + a_t \Delta t$ and $a_t$ is known for our body. The next section describes an even more accurate approach for $x_t$.
